---
title: "Abslute Value Constraints"
author: "Mark Newman"
output:
  html_document: default
editor_options: 
  chunk_output_type: console
---

# LP Formulation

Descision variables

* $X = x_{1}$

Objective function

* Maximize: $\lvert x_{1}-1 \rvert$
  
Subject to

* Constraint 1: $\lvert 3x_{1} \rvert <= 9$
* Bounds: $\infty <= x_{i} <= \infty \forall i$

---------

When $\lvert ax+n \rvert$ shows up in the objective function, convert it as follows:

Objective function

* $v = \lvert ax+n \rvert$

Subject to

* $ax-v <= -n$
* $-ax-v <= n$
* Bounds: $\0 <= v <= \infty$

When $\lvert ax+n \rvert <= c$ shows up a constraint, convert it as follows:

Subject to

* $ax-v <= -n$
* $-ax-v <= n$
* $v <= c$
* Bounds: $\0 <= v <= \infty$

Therefore, rewrite the LP Formulation

---------

# LP Formulation (rewriten)

Descision variables

* $X = x_{1}, v_{1}, v_{2}$

Objective function

* Maximize: $v_{1}$
    * $0x_{1} + v_{1} + 0v_{2}$
  
Subject to

* Objective function: $v_{1} = \lvert x_{1}-1 \rvert$
    * $x_{1}-v_{1} <= 1$
        * $x_{1} - v_{1} + 0v_{2} <= 1$
    * $-x_{1}-v_{1} <= -1$
        * $-x_{1} - v_{1} + 0v_{2} <= -1$
    * Bounds: $\0 <= v_{1} <= \infty$
* Constraint 1: $\lvert 3x_{1} \rvert <= 9$
    * $3x_{1}-v_{2} <= 0$
        * $3x_{1} + 0v_{1} - v_{2} <= 0$
    * $-3x_{1}-v_{2} <= 0$
        * $-3x_{1} + 0v_{1} - v_{2} <= 0$
    * $v_{2} <= 9$
        * $0x_{1} + 0v_{1} + v_{2} <= 9$
    * Bounds: $\0 <= v_{2} <= \infty$
* Bounds: $\infty <= x_{i} <= \infty$

# R translation

Clear everything
```{r}
rm(list = ls())
options(scipen = 999)
```

Required packages
```{r}
if (!require('lpSolveAPI')) install.packages('lpSolveAPI', quiet=TRUE)
library(lpSolveAPI)
```

Create model
```{r}
lpm = make.lp(ncol = 2, nrow = 0)
```

Name descision variables
```{r}
dimnames(lpm)[[2]] = c("x1", "v1")
```

Add objective function
```{r}
set.objfn(lpm, c(1, 0))
xxx = lp.control(lpm, sense = "max")
```

* `xxx` is there to suppress the output

Add in all constraints
```{r}
add.constraint(lpm, c(1,-1), "<=", 0)
add.constraint(lpm, c(-1,-1), "<=", 0)
add.constraint(lpm, c(0,1), "<=", 10)
add.constraint(lpm, c(1,0), "<=", 6)
set.bounds(lpm, lower = c(-50, 0), upper = c(50, 100))
#dimnames(lpm)[[1]] = c("obj 1a", "obj 1b")
```

max: |x-1|
st:
|x+2| <= 3

-->

http://cc.bingj.com/cache.aspx?q=lpsolveapi+absolute+value+objective+function&d=4801208246665729&mkt=en-US&setlang=en-US&w=ILTEw-NjHIeHVgwHCrI9_xR0_X1rHFsa

in the case of: max: +|ax+n| or min -|ax+n|

Obj
  |ax+n| => v

-ax +  v + -Ml <= n
 ax +  v +  Ml <= M-n
 ax + -v +  0l <= -n
-ax + -v +  0l <= n

for constraints

|ax+n| <= c

replaced with

 ax + -1x' <= -n
-ax + -1x' <= n
 0x +  1x' <= c

-->

max: v
st:
-1x +  0x' +  1v + -Ml <= -1
 1x +  0x' +  1v +  Ml <= M+1
 1x +  0x' + -1v +  0l <= 1
-1x +  0x' + -1v +  0l <= -1
 1x + -1x' +  0v +  0l <= -2
-1x + -1x' +  0v +  0l <= 2
 0x +  1x' +  0v +  0l <= 3

--> 

```{r}
rm(list = ls())
options(scipen = 999)
if (!require('lpSolveAPI')) install.packages('lpSolveAPI', quiet=TRUE)
library(lpSolveAPI)
lpm = make.lp(ncol = 4, nrow = 0)
dimnames(lpm)[[2]] = c("x", "x'", "v", "l")
set.objfn(lpm, c(0,0,1,0))
xxx = lp.control(lpm, sense = "max")
M = 100

add.constraint(lpm, c(-1, 0, 1,-M), "<=",  -1)
add.constraint(lpm, c( 1, 0, 1, M), "<=",  M+1)
add.constraint(lpm, c( 1, 0,-1, 0), "<=",  1)
add.constraint(lpm, c(-1, 0,-1, 0), "<=",  -1)
add.constraint(lpm, c( 1,-1, 0, 0), "<=",  -2)
add.constraint(lpm, c(-1,-1, 0, 0), "<=",  2)
add.constraint(lpm, c( 0, 1, 0, 0), "<=",  3)

set.type(lpm, c(4), "binary")
set.bounds(lpm, lower = c(-Inf,0,0,0))
res = solve.lpExtPtr(lpm)
res == 0
get.objective(lpm)
get.variables(lpm)
print.lpExtPtr(lpm)
delete.lp(lpm)
```



Solve
```{r results="hold"}
res = solve.lpExtPtr(lpm)
res == 0
get.objective(lpm)
get.variables(lpm)

print.lpExtPtr(lpm)
```

* `res == 0` is the status code for _optimal solution found_
* `get.objective(...)` gets the value of the _objective function_
* `get.variables(...)` gets the value of the _descision variables_

Cleanup
```{r}
delete.lp(lpm)
```

