---
title: "Abslute Value Constraints"
author: "Mark Newman"
output:
  html_document: default
editor_options: 
  chunk_output_type: console
---

# LP Formulation

Descision variables

* $X = x_{1}$

Objective function

* Maximize: $\lvert x-1 \rvert$
  
Subject to:

* Constraint 1: $x_{1} <= 3$
* Bounds: $\infty <= x_{i} <= \infty$

---------

When $\lvert x+n \rvert$ shows up in the objective function, convert it as follows:

* $\lvert x+n \rvert$
* $v >= x+n$
* $v >= -x-n$

Therefore, rewrite the LP Formulation

---------

# LP Formulation (rewriten)

Descision variables

* $X = x_{1}, v1_{1}$

Objective function

* Maximize: $v_{i}$
    * $0x_{1} + v_{1}$
  
Subject to:

* Constraint 1: $x_{1} <= 3$
    * $x_{1} + 0v_{1} <= 3$
* ABS 1a: $v_{1} >= x_{1}+1$
    * $0 >= x_{1} + 1 + -v_{1}$
    * $x_{1} + 1 + -v_{1} <= 0$
    * $x_{1} + -v_{1} <= -1$
* ABS 1b: $v_{1} >= -x_{1}-1$
    * $0 >= -x_{1} - 1 + -v_{1}$
    * -x_{1} - 1 + -v_{1} <= 0$
    * -x_{1} + -v_{1} <= 1$
* Bounds: $\infty <= x_{i}, v_{i} <= \infty$

# R translation

Clear everything
```{r}
rm(list = ls())
options(scipen = 999)
```

Required packages
```{r}
if (!require('lpSolveAPI')) install.packages('lpSolveAPI', quiet=TRUE)
library(lpSolveAPI)
```

Create model
```{r}
lpm = make.lp(ncol = 2, nrow = 0)
```

Name descision variables
```{r}
dimnames(lpm)[[2]] = c("x1", "v1")
```

Add objective function
```{r}
set.objfn(lpm, c(0, 1))
xxx = lp.control(lpm, sense = "max")
```

* `xxx` is there to suppress the output

Add in all constraints
```{r}
add.constraint(lpm, c(1, 0), "<=", 3)
add.constraint(lpm, c(1, -1), "<=", -1)
add.constraint(lpm, c(-1, -1), "<=", 1)
dimnames(lpm)[[1]] = c("Constraint 1", "ABS 1a", "ABS 1b" )
```

* ABS 1a: $v_{1} >= x_{1}+1$
    * $0 >= x_{1} + 1 + -v_{1}$
    * $x_{1} + 1 + -v_{1} <= 0$
    * $x_{1} + -v_{1} <= -1$
* ABS 1b: $v_{1} >= -x_{1}-1$
    * $0 >= -x_{1} - 1 + -v_{1}$
    * -x_{1} - 1 + -v_{1} <= 0$
    * -x_{1} + -v_{1} <= 1$
* Bounds: $\infty <= x_{i}, v_{i} <= \infty$

Solve
```{r results="hold"}
res = solve.lpExtPtr(lpm)
res == 0
get.objective(lpm)
get.variables(lpm)

print.lpExtPtr(lpm)
```

* `res == 0` is the status code for _optimal solution found_
* `get.objective(...)` gets the value of the _objective function_
* `get.variables(...)` gets the value of the _descision variables_

Cleanup
```{r}
delete.lp(lpm)
```

