---
title: "Basic Constraints"
author: "Mark Newman"
output:
  html_document: default
editor_options: 
  chunk_output_type: console
---

# LP Formulation

Descision variables

* $X = x_{1}$

Cost values

* $C = [2]$

Objective function

* Maximize: $\sum_{i=1}^{n} c_{i}x_{i}$
    * $2x_{1}$
  
Subject to:

* Constraint 1: $3x_{1} <= 9$
* Bounds: $0 <= x_{i} <= \infty$

# R translation

Clear everything
```{r}
rm(list = ls())
options(scipen = 999)
```

Required packages
```{r}
if (!require('lpSolveAPI')) install.packages('lpSolveAPI', quiet=TRUE)
library(lpSolveAPI)
```

Create model
```{r}
lpm = make.lp(ncol = 1, nrow = 0)
```

Name descision variables
```{r}
dimnames(lpm)[[2]] = c("x1")
```

Add objective function
```{r}
set.objfn(lpm, c(2))
xxx = lp.control(lpm, sense = "max")
```

* `xxx` is there to suppress the output

Add in all constraints
```{r}
add.constraint(lpm, c(3), "<=", 9)
set.bounds(lpm, lower = c(0))
dimnames(lpm)[[1]] = c("Constraint 1")
```

* `dimnames(...)...` is here to help make the `print.lpExtPtr(...)` command look nicer

Solve
```{r results="hold"}
res = solve.lpExtPtr(lpm)
res == 0
get.objective(lpm)
get.variables(lpm)

print.lpExtPtr(lpm)
```

* `res == 0` is the status code for _optimal solution found_
* `get.objective(...)` gets the value of the _objective function_
* `get.variables(...)` gets the value of the _descision variables_

Cleanup
```{r}
delete.lp(lpm)
```

