---
title: "Linked Constraints"
author: "Mark Newman"
output:
  html_document: default
editor_options: 
  chunk_output_type: console
---

# LP Formulation

Notes on **conventions**

* When talking about descision variables and constraints, it is convention to list out the real variable ($x_{i}$) then linking variable ($y_{i}$)
* When talking about costs and the objective function, it is convention to list out fixed costs ($f_{i}$) then variable costs ($c_{i}$)

What this means is that you need to be careful to flip one or the other when translating it to R. As seen below, I prefer to flip the objective function and put all my $x_{i}$s up front.


Descision variables

* $X = x_{1}, y_{1}$

Costs

* $F = [-50]$
* $C = [2]$

Objective function

* Maximize: Fixed costs + Variable costs
    * Fixed costs: $\sum_{i=1}^{n} f_{i}y_{i}$
        * $-50y_{1}$
    * Variable costs: $\sum_{i=1}^{n} c_{i}x_{i}$
        * $2x_{1}$
    * $-50y_{1} + 2x_{1}$
    * $2x_{1} + -50y_{1}$
  
Subject to:

* Constraint 1: $3x_{1} <= 200$
    * $3x_{1} + 0y_{1} <= 200$
* Links: $x_{i} <= My_{i}$
    * $x_{i} + -My_{i} <= 0$
    * $x_{1} + -My_{1} <= 0$
        * M needs to be larger than any posible $x_{i}$ so choose the bigest constraint + 1.
    * $x_{1} + -201y_{1} <= 0$
* Binary: $y_{i}$
* Bounds: $0 <= x_{i} <= \infty$

# R translation

Clear everything
```{r}
rm(list = ls())
options(scipen = 999)
```

Required packages
```{r}
if (!require('lpSolveAPI')) install.packages('lpSolveAPI', quiet=TRUE)
library(lpSolveAPI)
```

Create model
```{r}
lpm = make.lp(ncol = 2, nrow = 0)
```

Name descision variables
```{r}
dimnames(lpm)[[2]] = c("x1", "y1")
```

Add objective function
```{r}
set.objfn(lpm, c(2, -50))
xxx = lp.control(lpm, sense = "max")
```

* `xxx` is there to suppress the output

Add in all constraints
```{r}
add.constraint(lpm, c(3, 0), "<=", 200)
add.constraint(lpm, c(1, -201), "<=", 0)
set.type(lpm, c(2), "binary")
set.bounds(lpm, lower = c(0,0))
dimnames(lpm)[[1]] = c("Constraint 1", "Link x1")
```

* `dimnames(...)...` is here to help make the `print.lpExtPtr(...)` command look nicer

Solve
```{r results="hold"}
res = solve.lpExtPtr(lpm)
res == 0
get.objective(lpm)
get.variables(lpm)

print.lpExtPtr(lpm)
```

* `res == 0` is the status code for _optimal solution found_
* `get.objective(...)` gets the value of the _objective function_
* `get.variables(...)` gets the value of the _descision variables_

Cleanup
```{r}
delete.lp(lpm)
```

