---
title: "11.3.1"
author: "JMark Newman"
output:
  html_document: default
editor_options: 
  chunk_output_type: console
---

# Clear everything
```{r}
rm(list = ls())
options(scipen = 999)
```

# Required packages
```{r message=FALSE, warning=FALSE}
if (!require('lpSolve')) install.packages('lpSolve', quiet=TRUE)
```

# LP Formulation

Descision variables

* p1, p2, p3, p4, p5

Maximize:

* 10p1 + 17p2 + 16p3 + 8p4 + 14p5
  
Subject to:

* 48p1 + 96p2 + 80p3 + 32p4 + 64p5 <= 160

Objective function
```{r}
obj = c(10, 17, 16, 8, 14)
```

Requirments
```{r}
req.capital = c(48, 96, 80, 32, 64)
```

* `lp` assumes varables >=0

Constraints
```{r}
con.capital = 160
```

Solve (general LP)
```{r}
lhs = rbind(req.capital)
dir = c("<=")
rhs = c(con.capital)


model = lp ("max", obj, lhs, dir, rhs)
model$objval
model$solution

cbind((rowSums(lhs*model$solution)), dir, rhs)
```

* `model$objval` is the value of the __Objective Function__
* `model$solution` are the values of the __Descision Variables__
* `cbind(...)` is there so you can _double check_ your work

Solve (binary LP)
```{r}
model = lp ("max", obj, lhs, dir, rhs, all.bin = TRUE)
model$objval
model$solution

cbind((rowSums(lhs*model$solution)), dir, rhs)
```

* The only thing that changed was in `all.int = TRUE` to `all.bin = TRUE`

