---
title: "11.4.2 Linking Constraints and Fixed Costs"
author: "Mark Newman"
output:
  html_document: default
---

# Clear everything
```{r}
rm(list = ls())
options(scipen = 999)
```

# Required packages
```{r message=FALSE, warning=FALSE}
if (!require('lpSolve')) install.packages('lpSolve', quiet=TRUE)
source("../Helper Scripts/rbindPattern.r")
source("../Helper Scripts/lpPrettyPrint.r")
```

# LP Formulation (no fixed costs)

Descision variables

* F1, F2, F3

Maximize:

* 1.2F1 + 1.8F2 + 2.2F3
  
Subject to:

* Department A: 3F1 + 4F2 + 8F3 <= 2000
* Department B: 3F1 + 5F2 + 6F3 <= 2000
* Department C: 2F1 + 3F2 + 9F3 <= 2000
* Demand:
    * F1 <= 400
    * F2 <= 300
    * F3 <= 50
  
# R translation

Descision variables
```{r}
dv = c("F1","F2","F3")
```

Objective function
```{r}
obj = c(1.2, 1.8, 2.2)
names(obj) = dv
```

Requirments
```{r}
req.depa = c(3,4,8)
req.depb = c(3,5,6)
req.depc = c(2,3,9)
req.df1 = c(1,0,0)
req.df2 = c(0,1,0)
req.df3 = c(0,0,1)
```

* `lp` assumes varables >=0

Constraints
```{r}
con.depa = 2000
con.depb = 2000
con.depc = 2000
con.df1 = 400
con.df2 = 300
con.df3 = 50
```

Solve
```{r}
lhs = rbindPattern("req\\..+")
dir = rep_len("<=", dim(lhs)[1])
rhs = rbindPattern("con\\..+")

model = lpSolve::lp("max", obj, lhs, dir, rhs, all.int = TRUE)
lppp(model)
```

* The regular expression `req\\..+` matches anything that starts with "req." and has at least one other character on the end
* `lppp(...)` so you can see a pretty print of the `model`

# Book results

Per page 320 "The optimal product mix ... F1 at 400k, F2 at 100K, and F3 at 50k. This product mix creates $700K in variable profits."

This matches our result `$Solution` F1=400,F2=100,F3=50 yealds 770

# LP Formulation (with linked fixed costs)

Descision variables

* F1, F2, F3, LF1, LF2, LF3

Maximize:

* 1.2F1 + 1.8F2 + 2.2F3 - 60LF1 - 200LF2 - 100LF3
  
Subject to:

* Department A: 3F1 + 4F2 + 8F3 <= 2000
* Department B: 3F1 + 5F2 + 6F3 <= 2000
* Department C: 2F1 + 3F2 + 9F3 <= 2000
* Demand:
    * F1 <= 400
    * F2 <= 300
    * F3 <= 50
* Links:
    * F1 - MLF1 <= 0 
    * F2 - MLF2 <= 0 
    * F3 - MLF3 <= 0 

In the "_Links_" section the form is

* x <= M*y
* x - My <= 0

where

* x is continous profit
* y is the linked boolean corsponding to cost
* M is very large. See below at to what M is.

# R translation

Descision variables
```{r}
dv = c("F1","F2","F3","LF1","LF2","LF3")
```

Objective function
```{r}
obj = c(1.2, 1.8, 2.2, -60, -200, -100)
names(obj) = dv
```

Requirments
```{r}
req.depa = c(3,4,8,0,0,0)
req.depb = c(3,5,6,0,0,0)
req.depc = c(2,3,9,0,0,0)
req.df1 = c(1,0,0,0,0,0)
req.df2 = c(0,1,0,0,0,0)
req.df3 = c(0,0,1,0,0,0)
req.lf1 = c(1,0,0,-401,0,0)
req.lf2 = c(0,1,0,0,-401,0)
req.lf3 = c(0,0,1,0,0,-401)
```

* We added 3 linking variables that means we need to add `,0,0,0` to the end of all existing requirments
* The `-401` comes from M. It was chosen because the max(con.df#) = 400 < 401

Constraints
```{r}
con.depa = 2000
con.depb = 2000
con.depc = 2000
con.df1 = 400
con.df2 = 300
con.df3 = 50
con.lf1 = 0
con.lf2 = 0
con.lf3 = 0
```

Solve
```{r}
lhs = rbindPattern("req\\..+")
dir = rep_len("<=", dim(lhs)[1])
rhs = rbindPattern("con\\..+")

model = lpSolve::lp("max", obj, lhs, dir, rhs, int.vec = c(1,2,3), binary.vec = c(4,5,6))
lppp(model)
```

* `int.vec` and `binary.vec` lets solver sort out what the variable types are

# Book results

Per page 322 "The optimal solution ...profit of \$508k. ... F1 up to its ceiling(400) ... F2 at the level of 160K ..."

This matches our result `$Solution` with 0s and binary links removed F1=400,F2=160 yealds 508