---
title: "11.5.1 Facility Location Model (Capacitated)"
author: "Mark Newman"
output:
  html_document: default
editor_options: 
  chunk_output_type: console
---

# Clear everything
```{r}
rm(list = ls())
options(scipen = 999)
```

# Required packages
```{r message=FALSE, warning=FALSE}
if (!require('lpSolve')) install.packages('lpSolve', quiet=TRUE)
source("../Helper Scripts/lpPrettyPrint.r")
source("../Helper Scripts/transportation.r")
```

# Tables

## From

From | Capacity | Cost
:--:|:--:|:--:
(A)lbuquerque | 16 | 140
(Da)llas | 20 | 150
(De)nver | 10 | 100
(H)ouston | 10 | 110
(P)hoenix | 12 | 125
(S)an (A)ntonio | 12 | 120

## To

To | Demand
:--:|:--:
(A)lbuquerque | 3.2
(B)oise | 2.5
(Da)llas | 6.8
(De)nver | 4.0
(H)ouston | 9.6
(O)klahoma | 3.5
(P)hoenix | 5.0
(S)alt (L)ake (C)ity | 1.8
(S)an (A)ntonio | 7.4
(W)ichita | 2.7

## Cost
From \\ To | A | B | Da | De | H | O | P | SLC | SA | W
:--:|:--:|:--:|:--:|:--:|:--:|:--:|:--:|:--:|:--:|:--:
A  | 00.00 | 47.00 | 32.00 | 22.00 | 42.50 | 27.00 | 23.00 | 30.00 | 36.50 | 29.50
Da | 32.00 | 79.50 | 00.00 | 39.00 | 12.50 | 10.50 | 50.00 | 63.00 | 13.50 | 17.00
De | 21.00 | 42.00 | 39.00 | 00.00 | 51.50 | 31.50 | 40.50 | 24.00 | 47.50 | 26.00
H  | 42.50 | 91.00 | 12.50 | 51.50 | 00.00 | 23.00 | 58.00 | 72.00 | 10.00 | 31.00
P  | 23.00 | 49.00 | 50.00 | 40.50 | 58.00 | 49.00 | 00.00 | 32.50 | 50.00 | 52.00
SA | 36.50 | 83.50 | 13.50 | 47.50 | 10.00 | 24.00 | 50.00 | 66.50 | 00.00 | 32.00

# LP Formulation

Descision variables

* xij = quanity shiped from i to j
    * i = A, Da, De, H, P, SA
    * j = A, B, Da, De, H, O, P, SLC, SA, W
* yi = is location i open

Minimize:

* Fixed Cost + Variable Cost
    * Fixed Cost: sum(i\*costi)
    * Variable Cost: sum(xij\*costij)

Subject to:

* yi is binary
* sent <= capacity
* demand <= received
* Links:
    * xij - Myi <= 0

# R translation

Tables
```{r}
from = data.frame(matrix(c("A",16,140,"Da",20,150,"De",10,100,"H",10,110,"P",12,125,"SA",12,120), nrow = 6, ncol=3, byrow=TRUE), stringsAsFactors = FALSE)
colnames(from) = c("From", "Capacity", "Cost")
from$Capacity = as.numeric(from$Capacity)
from$Cost = as.numeric(from$Cost)

to = data.frame(matrix(c("A",3.2,"B",2.5,"Da",6.8,"De",4.0,"H",9.6,"O",3.5,"P",5.0,"SLC",1.8,"SA",7.4,"W",2.7), nrow = 10, ncol=2, byrow=TRUE), stringsAsFactors = FALSE)
colnames(to) = c("To", "Demand")
to$Demand = as.numeric(to$Demand)

cost = data.frame(matrix(c(00.00,47.00,32.00,22.00,42.50,27.00,23.00,30.00,36.50,29.50,32.00,79.50,00.00,39.00,12.50,10.50,50.00,63.00,13.50,17.00,21.00,42.00,39.00,00.00,51.50,31.50,40.50,24.00,47.50,26.00,42.50,91.00,12.50,51.50,00.00,23.00,58.00,72.00,10.00,31.00,23.00,49.00,50.00,40.50,58.00,49.00,00.00,32.50,50.00,52.00,36.50,83.50,13.50,47.50,10.00,24.00,50.00,66.50,00.00,32.00), nrow = 6, ncol=10, byrow=TRUE), stringsAsFactors = FALSE)
colnames(cost) = to$To
rownames(cost) = from$From
```

Descision variables
```{r}
dv.fixed = sprintf("lnk-%s", from$From)
dv.variable = fromTo2DV(from$From, to$To)
dv = c(dv.fixed, dv.variable)

rm(dv.fixed, dv.variable)
```

* `sprintf(...)` is used because we want to be clear on the 2 different types of DVs 

Objective function
```{r}
obj.fixed = from$Cost
obj.variable = as.vector(t(cost))
obj = c(obj.fixed, obj.variable)
names(obj) = dv

rm(obj.fixed, obj.variable)
length(dv) == length(obj)
```

* naming (`names(obj) = dv`) our objective function variables makes `lppp(...)` look nicer
* `length(dv) == length(obj)` is here because we want to double check our work

Requirments
```{r}
lenf = length(from$Capacity)
lent = length(to$Demand)

req.capacity = CapacityRequirments(lenf, lent)
req.demand =  DemandRequirments(lenf, lent)
req.link = LinkRequirments("from", lenf, lent, sum(to$Demand)+1)

req.capacity = cbind(matrix(0, nrow = lenf, ncol = lenf), req.capacity)
req.demand = cbind(matrix(0, nrow = lent, ncol = lenf), req.demand)

req = rbind(req.capacity, req.demand, req.link)
rm(lenf, lent, req.capacity, req.demand, req.link)
length(dv) == dim(req)[2]
```

* `lp` assumes varables >=0
* The form of the capacity requirment for 2 "from"s and 3 "to"s is
    * f1: c(1,1,1,0,0,0)
    * f2: c(0,0,0,1,1,1)
* The form of the demand requirment for 2 "from"s and 3 "to"s is
    * t1: c(1,0,0,1,0,0)
    * t2: c(0,1,0,0,1,0)
    * t3: c(0,0,1,0,0,1)
* When the link is on the "from", the form of the link requirment for 2 "from"s and 3 "to"s is
    * l1: c(-M,0,1,1,1,0,0,0)
    * l2: c(0,-M,0,0,0,1,1,1)
    * M needs to be set _big_ usualy I set it to sum(demand) + 1. That is because pulling _anything_ from the "from" means you could pull _everything_ from there. The max pull is sum(demand).
* Notice the up front padding (`cbind(matrix(...)], ...)`). It is there because we need a place for our linked variables. `CapacityRequirments` and `DemandRequirments` are based on the general transport model so we need to tack on some room.
    

Constraints
```{r}
con.capacity = from$Capacity
con.demand = to$Demand
con.link = rep_len(0, length(from$Capacity))
con = c(con.capacity, con.demand, con.link)

rm(con.capacity, con.demand, con.link)
length(con) == dim(req)[1]
```

* `con.link` is always 0. We just need to put in enough of them 

Solve
```{r}
lenf = length(from$Capacity)
lent = length(to$Demand)

dir.capacity = rep_len("<=", lenf)
dir.demand = rep_len(">=", lent)
dir.link = rep_len("<=", lenf)
dir = c(dir.capacity, dir.demand, dir.link)

model = lpSolve::lp("min", obj, req, dir, con, binary.vec=1:lenf)
lppp(model)
rm(dir.capacity, dir.demand, dir.link)
```

* `lppp(...)` so you can see a pretty print of the `model`

# Book results

Per page 327 "... shown in Figure11.21, which achieves a minimum total cost of $884,550."

* Links: Da,De,H,P
* Da: Da=6800,0=3500,SA=7000,W=2700
* De: A=1700,B=2500,De=4000,SLC=1800
* H: H=9600,SA=400
* P: A=1500,P=5000

This matches our result `$Solution` when scaled by 1k

* lnk-A=0,lnk-Da=1,lnk-De=1,lnk-H=1,lnk-P=1,lnk-SA=0,AA=0,AB=0,ADa=0,ADe=0,AH=0,AO=0,AP=0,ASLC=0,ASA=0,AW=0,DaA=0,DaB=0,DaDa=6.8,DaDe=0,DaH=0,DaO=3.5,DaP=0,DaSLC=0,DaSA=7,DaW=2.7,DeA=1.7,DeB=2.5,DeDa=0,DeDe=4,DeH=0,DeO=0,DeP=0,DeSLC=1.8,DeSA=0,DeW=0,HA=0,HB=0,HDa=0,HDe=0,HH=9.6,HO=0,HP=0,HSLC=0,HSA=0.400000000000001,HW=0,PA=1.5,PB=0,PDa=0,PDe=0,PH=0,PO=0,PP=5,PSLC=0,PSA=0,PW=0,SAA=0,SAB=0,SADa=0,SADe=0,SAH=0,SAO=0,SAP=0,SASLC=0,SASA=0,SAW=0 yealds 884.55
* lnk-Da=1,lnk-De=1,lnk-H=1,lnk-P=1,DaDa=6.8,DaO=3.5,DaSA=7,DaW=2.7,DeA=1.7,DeB=2.5,DeDe=4,DeSLC=1.8,HH=9.6,HSA=0.4,PA=1.5,PP=5yealds 884.55
    * links: Da,De,H,P
    * Da: Da=6.8,O=3.5,SA=7,W=2.7
    * De: A=1.7,B=2.5,De=4,SLC=1.8
    * H: H=9.6,SA=0.4
    * P: A=1.5,P=5

