---
title: "10.2 Transportation Model"
author: "Mark Newman"
output:
  html_document: default
---

# Clear everything
```{r}
rm(list = ls())
options(scipen = 999)
```

# Required packages
```{r message=FALSE, warning=FALSE}
if (!require('lpSolve')) install.packages('lpSolve', quiet=TRUE)
source("../Helper Scripts/rbindPattern.r")
source("../Helper Scripts/lpPrettyPrint.r")
```

# LP Formulation

Descision variables

* From: (M)inneapolis, (P)ittsburgh, (T)ucson
* To: (A)tlanta, (B)oston, (C)hicago, (D)enver

Objective function

* Minimize: .6MA + .56MB + .22MC + .4MD + .36PA + .3PB + .28PC + .58PD + .65TA + .68TB + .55TC + .42TD
  
Subject to:

* Minneapolis: MA + MB + MC + MD <= 9000
* Pittsburgh: PA + PB + PC + PD <= 12000
* Tucson: TA + TB + TC + TD <= 13000
* Atlanta: MA + PA + TA >= 7500
* Boston: MB + PB + TB >= 8500
* Chicago: MC + PC + TC >= 9500
* Denver: MD + PD + TD >= 8000

# R translation

Descision variables
```{r}
dv.f = c("m", "p", "t")
dv.t = c("a", "b", "c", "d")
dv = c("MA","MB", "MC", "MD", "PA", "PB", "PC", "PD", "TA", "TB", "TC", "TD")
```

* the `dv` here is the cross product of the to and from variables

Cost Matrix
```{r}
cost.m = c(.60, .56, .22, .40)
cost.p = c(.36, .30, .28, .58)
cost.t = c(.65, .68, .55, .42)
cost = rbindPattern("cost\\..+")
```

* The regular expression `cost\\..+` matches anything that starts with "cost." and has at least one other character on the end
* Format your costs with the same number of decimal places as a form of visual QA

Capacity & Demand
```{r}
cap = c(9000, 12000, 13000)
dem = c(7500, 8500, 9500, 8000)
```

Objective function
```{r}
obj = as.vector(t(cost))
names(obj) = dv
```

* `as.vector(t(cost))` unrolls the cost matrix above by row, which matched how we declaired our `dv`
* naming (`names(obj) = dv`) our objective function variables makes `lppp(...)` look nicer

Requirments
```{r}
req.1m = c(1, 1, 1, 1, 0, 0, 0, 0, 0, 0, 0, 0)
req.2p = c(0, 0, 0, 0, 1, 1, 1, 1, 0, 0, 0, 0)
req.3t = c(0, 0, 0, 0, 0, 0, 0, 0, 1, 1, 1, 1)
req.4a = c(1, 0, 0, 0, 1, 0, 0, 0, 1, 0, 0, 0)
req.5b = c(0, 1, 0, 0, 0, 1, 0, 0, 0, 1, 0, 0)
req.6c = c(0, 0, 1, 0, 0, 0, 1, 0, 0, 0, 1, 0)
req.7d = c(0, 0, 0, 1, 0, 0, 0, 1, 0, 0, 0, 1)
```

* `lp` assumes all varables are >= 0, so we don't need to add that in too
* `req.#??` is a special naming convention used by `rbindPattern` which will line up the left hand side (`lhs`) and right hand side (`rhs`) for us in an easy manner
  * The # in `req.#??` is to force the sort order so we can make sure we get the direction (>= & <=) in the right order

Constraints
```{r}
con = c(cap, dem)
```

Solve
```{r}
lhs = rbindPattern("req\\..+")
dir = c("<=", "<=", "<=", ">=", ">=", ">=", ">=")
rhs = con

model = lpSolve::lp ("min", obj, lhs, dir, rhs, all.int=TRUE)
model$objval
model$solution

lppp(model)
```

* `all.int=TRUE` forces in integer solution. If you dont have this TC will be 499.999999999999
* `model$objval` is the value of the __Objective Function__
* `model$solution` are the values of the __Descision Variables__
* `lppp(...)` so you can see a pretty print of the `model`


# Book results

Per page 280 "... the solution shown in Figure 10.4, which achieves the minimum cost of $12,025"

* MC = 9000
* PA = 3500, PB = 8500
* TA = 4000, TC = 500, TD = 8000 

This matches our result `$Solution` MA=0,MB=0,MC=9000,MD=0,PA=3500,PB=8500,PC=0,PD=0,TA=4000,TB=0,TC=500,TD=8000 yealds 12025

* Remove 0s: MC=9000,PA=3500,PB=8500,TA=4000,TC=500,TD=8000 yealds 12025



