---
title: "12.7 Machine-Sequencing Problem (2 machine)"
author: "Mark Newman"
output:
  html_document: default
editor_options: 
  chunk_output_type: console
---

The purpose of this problem is to extend the earler 1 machine sample into 2 machines. This is not in the book so I can't know the best solution

# Tables

|Job Number | 1 | 2 | 3 | 4 | 5 | 6 |
|:--:|:--:|:--:|:--:|:--:|:--:|:--:|
| Processing time M1 | 15 | 17 | 19 | 21 | 23 | 25 |
| Processing time M2 | 16 | 18 | 20 | 20 | 22 | 24 |
| Due date | 28 | 35 | 24 | 32 | 30 | 40 |

# Formulation

Static variables

$D = d_{x}$

* The due date of job $x$
    * $D = [28,35,24,32,30,40]$
    
$P = p_{x}$

* The time to process job $x$
    * $P = [5,7,9,11,13,15]$

Descision variables

$X = x_{1}, x_{2}, ... x_{6}, y_{1}$

* $x_{i}$ is the job number in position x.
    * Processing job 2 in position 4 would mean $x_{4} = 2$
* $y_{1}$ is the number of jobs to be processed on machine x

Objective function

* Minimize: $\max(X_{1}, X_{2})$
    * $X_{1} = \sum_{i=1}^{y_{1}}\max(0, \sum_{j=1}^{i}p_{x_{j}} - d_{x_{j}})$
    * $X_{2} = \sum_{i=(y_{1}+1)}^{6}\max(0, \sum_{j=1}^{i}p_{x_{j}} - d_{x_{j}})$

Subject to

* All different $x_{i}$ 
* Bounds
     * $1 <= x_{i} <= 6$
     * $0 <= y_{1} <= 6$

# R translation

**Process**

* Clear everything
* Required packages
* Setup tables
* Add fitness function
    * `JDEoptim(...)` only supports real valued inputs. To _force_ some of those to be ints use `floor(...)` in the objective function
    * Test `fn(...)`
* Add in the constraints
    * All constraints should be re-written in the form of $f(x) = 0$ or $f(x) <= 0$
    * `JDEoptim(...)` only supports real valued inputs. To _force_ some of those to be ints use `floor(...)` in the objective function
    * In the case of straight `TRUE`/`FALSE` constraints like "All Different" return `-1` and make that an equality constraint. If you make it `+1`,this algorithm freaks out and stops. I don't know why.
    * Test `constr(...)`
* Add in bounds
* Use `tol = 0.000001`
    * The default tolarance is _too_ small for most cases
* Use `set.seed(...)`
    * This ensures predictable results when re-running.
* Solve
    * `JDEoptim(...)` only supports real valued inputs. To _allow_ some of those to be ints, use `(ub+1)`
    * `meq=1` forces the first $1$ results from `constr(...)` to be considered as $=0$ then the rest as $<=0$
    * `model$value` is the result of the objective function.
        * $Objective = 29$
    * `model$par` are the values of the desision variables that made the `model$value`
        * $X = [3,1,4,5,2,6,3]$
* Double check all constraints are meet
* Solve (exhaust)
    * This is a sample that is not part of the book. Therefor double check the only way there is, exhaust the search space.

**R Results**

The minimum over due is $29$ using the sequence $X = [3,1,4,5,2,6,3]$

* Machine 1 = $X_{1} = [3,1,4]$
* Machine 2 = $X_{2} = [5,2,6]$

**R results (exhaust)**

The minimum over due is $29$ using the sequence $X = [3,1,4,5,2,6,3]$ based on an exhaustive search

* Machine 1 = $X_{1} = [3,1,4]$
* Machine 2 = $X_{2} = [5,2,6]$

**Code**
```{r results="hold"}
rm(list = ls())
options(scipen = 999)
if (!require('DEoptimR')) install.packages('DEoptimR', quiet=TRUE)
library(DEoptimR)
tab = matrix(c(15,17,19,21,23,25,16,18,20,20,22,24,28,35,24,32,30,40), nrow = 3, ncol = 6, byrow = TRUE)
fn = function(x) {
  
  x = floor(x)
  seq = x[1:(length(x)-1)]
  m1 = x[length(x)] 
  m2 = length(seq) - m1 

  fitness = c(0,0)
  if(m1 > 0) {
    
    s = seq[1:m1]
    pro = tab[1, s] # processing time re-ordered by sequence
    dro = tab[3, s] # due date re-ordered by sequence
    
    for(i in 1:m1) {
      fitness[1] = fitness[1] + max(0, sum(pro[1:i]) - dro[i])
    }
  }
  if(m2 > 0) {
    s = seq[(m1+1):(m2+m1)]
    pro = tab[2, s] # processing time re-ordered by sequence
    dro = tab[3, s] # due date re-ordered by sequence
    
    for(i in 1:m2) {
      fitness[2] = fitness[2] + max(0, sum(pro[1:i]) - dro[i])
    }
  }

  return (max(fitness))
}
constr = function(x) {
  
  x = floor(x)
  seq = x[1:(length(x)-1)]
  constraints = c(0)
  
  if(length(unique(seq)) != length(seq))
    constraints[1] = -1

  return(constraints)
}
fn(c(1:6,6)) == 212
fn(c(1:6,0)) == 218
fn(c(1:6,1)) == 154
constr(c(1,2,3,4,5,5,1)) == c(-1)
constr(c(1,2,3,4,5,6,1)) == c(0)
lb = c(1,1,1,1,1,1,0)
ub = c(6,6,6,6,6,6,6)
tol = 0.000001
set.seed(1)
model = JDEoptim(lower=lb, upper=(ub+1), fn=fn, constr=constr, meq=1, tol=tol)
model$value
floor(model$par)
all(constr(floor(model$par)) == 0)
```

**Code (exhaust)**
```{r results="hold"}
if (!require('gtools')) install.packages('gtools', quiet=TRUE)
library(gtools)
vals = permutations(n=6, r=6, v=1:6)
besti = 0
bestj = 0
bestfn = 10000
for(i in 1:nrow(vals)) {
  for(j in 0:6) {
    t1 = c(vals[i,], j)
    if(constr(t1) == c(0)) {
      currfn = fn(t1)
      if(currfn < bestfn) {
        bestfn = currfn
        besti = i
        bestj = j
      }
    }
  }
}
bestfn
c(vals[besti,], bestj)
```
