---
title: "12.7 Machine-Sequencing Problem (2 machine)"
author: "Mark Newman"
output:
  html_document: default
editor_options: 
  chunk_output_type: console
---

The purpose of this problem is to extend the earler 1 machine sample into 2 machines. This is not in the book so I can't know the best solution

# Tables

|Job Number | 1 | 2 | 3 | 4 | 5 | 6 |
|:--:|:--:|:--:|:--:|:--:|:--:|:--:|
| Processing time M1 | 15 | 17 | 19 | 21 | 23 | 25 |
| Processing time M2 | 16 | 18 | 20 | 20 | 22 | 24 |
| Due date | 28 | 35 | 24 | 32 | 30 | 40 |

# Formulation

Static variables

$D = d_{x}$

* The due date of job $x$
    * $D = [28,35,24,32,30,40]$
    
$P = p_{x}$

* The time to process job $x$
    * $P = [5,7,9,11,13,15]$

Descision variables

$X = x_{1}, x_{2}, ... x_{6}, y_{1}$

* $x_{i}$ is the job number in position x.
    * Processing job 2 in position 4 would mean $x_{4} = 2$
* $y_{1}$ is the number of jobs to be processed on machine x

Objective function

* Minimize: $\max(X_{1}, X_{2})$
    * $X_{1} = \sum_{i=1}^{y_{1}}\max(0, \sum_{j=1}^{i}p_{x_{j}} - d_{x_{j}})$
    * $X_{2} = \sum_{i=(y_{1}+1)}^{6}\max(0, \sum_{j=1}^{i}p_{x_{j}} - d_{x_{j}})$

Subject to

* All different $x_{i}$ 
* Bounds
     * $1 <= x_{i} <= 6$
     * $0 <= y_{1} <= 6$

# R translation

**Process**

* Clear everything
* Required packages
* Setup tables
* When the objective is to minimize, GA constraint violations need to force large $M$ for the fitness function. Assume the worst for everything, then double.
* Add fitness function
    * `genoud(...)` apperes to only allows for the notion of "fitness" and not constraints. So calculate the `fitness`, then evaluate all the `constraints`. If any of the constraints are violated force $M$
        * Objective Min should update the result to a large value
    * Test `fn(...)`
        * `fn(1:6,6)` means process all the work on machiene #1 and leave the other idle
* Add in the Domain. The first column is the lower bound and the second column is the upper bound.
* Solve
    * When using a genetic algorithm solver, always do at least 2 starting positions. In-order 50/50 (`1:6,3`) and reverse-order 50/50 (`6:1,3`) seem like good bets.
    * Use `set.seed(...)` to get predictable results when re-running.
    * Use `data.type.int = TRUE` whenever posible else the GA will tend to return a float
    * Use `Domains` but **NOT** `boundary.enforcement=2`. This may seem counter intituive, but if the generation can't "grow" outside the space, it has a hard time aproaching the line.
    * Use `print.level = 0` because `genoud(...)` is _super_ chatty otherwise.
    * `$value`  is the result of the objective function.
    * `$par` are the values of the desision variables that made the `$value`
* Solve (exhaust)
    * This is a sample that is not part of the book. Therefor double check the only way there is, exhaust the search space.

**R Results**

The minimum over due is $54$ using the sequence $X = [3,2,5,1,4,6,3]$ based on `model1`

* Machine 1 = $X_{1} = [3,2,5]$
* Machine 2 = $X_{2} = [1,4,6]$
* `model2` performs the same, but with a different sequence $X = [3,2,6,1,4,5,3]$

**R results (exhaust)**

The minimum over due is $54$ using the sequence $X = [3,2,5,1,4,6,3]$ based on an exhaustive search

* Machine 1 = $X_{1} = [3,2,5]$
* Machine 2 = $X_{2} = [1,4,6]$

**Code**
```{r results="hold"}
rm(list = ls())
options(scipen = 999)
if (!require('DEoptimR')) install.packages('DEoptimR', quiet=TRUE)
library(DEoptimR)
tab = matrix(c(15,17,19,21,23,25,16,18,20,20,22,24,28,35,24,32,30,40), nrow = 3, ncol = 6, byrow = TRUE)
fn = function(x) {
  
  x = floor(x)
  seq = x[1:(length(x)-1)]
  m1 = x[length(x)] 
  m2 = length(seq) - m1 

  fitness = 0
  if(m1 > 0) {
    
    s = seq[1:m1]
    pro = tab[1, s] # processing time re-ordered by sequence
    dro = tab[3, s] # due date re-ordered by sequence
    
    for(i in 1:m1) {
      fitness = fitness + max(0, sum(pro[1:i]) - dro[i])
    }
  }
  if(m2 > 0) {
    s = seq[(m1+1):(m2+m1)]
    pro = tab[2, s] # processing time re-ordered by sequence
    dro = tab[3, s] # due date re-ordered by sequence
    
    for(i in 1:m2) {
      fitness = fitness + max(0, sum(pro[1:i]) - dro[i])
    }
  }
  
  if(length(unique(seq)) != length(seq))
    fitness = M

  return (fitness)
}
constr = function(x) {
  
  x = floor(x)

    
  return(constraints)
}
#fn(c(1:6,6)) == 212
#fn(c(1:6,0)) == 218
#fn(c(1:6,1)) == 154
#fn(c(1,2,3,4,5,5,6)) == M
lb = c(1,1,1,1,1,1,1)
ub = c(6,6,6,6,6,6,5)
#sp1 = c(1:6,3)
#sp2 = c(6:1,3)
#set.seed(1)
#model1 = genoud(fn=fn, nvars=7, starting.values=sp1, Domains=Domains, print.level = 0, data.type.int = TRUE)
#set.seed(1)
#model2 = genoud(fn=fn, nvars=7, starting.values=sp2, Domains=Domains, print.level = 0, data.type.int = TRUE)
#model1$value
#model2$value
#model1$par
#model2$par
```

**Code (exhaust)**
```{r results="hold"}
#if (!require('gtools')) install.packages('gtools', quiet=TRUE)
#library(gtools)
#vals = permutations(n=6, r=6, v=1:6)
#besti = 0
#bestj = 0
#bestfn = M
#for(i in 1:nrow(vals)) {
#  for(j in 0:6) {
#    currfn = fn(c(vals[i,], j))
#    if(currfn < bestfn) {
#      bestfn = currfn
#      besti = i
#      bestj = j
#    }
#  }
#}
#bestfn
#c(vals[besti,], bestj)
```
