---
title: "9.2 Allocation Models (nonlinear approach)"
author: "Mark Newman"
output:
  html_document: default
editor_options: 
  chunk_output_type: console
---

The intent of this extended example is to look at the difference between the clasic lp formulation and the nonlinear version to see how the different R packages work. To this end the format will be different starting with the result then looking at the 2 different methods. Read this **AFTER** "9.2 Allocation Models".

# Clear everything
```{r}
rm(list = ls())
options(scipen = 999)
```

# Required packages
```{r message=FALSE, warning=FALSE}
if (!require('lpSolve')) install.packages('lpSolve', quiet=TRUE)
if (!require('NlcOptim')) install.packages('NlcOptim', quiet=TRUE)
source("../Helper Scripts/basicBindings.r")
source("../Helper Scripts/rbindPattern.r")
```

* `lpSolve` is one of the best linear programing packages for R. Use this as a baseline.
* `NlcOptim` is a popular R package implementing a nonlinear solver.

# Book results

Per page 245 "As shown in Figure 9.6, ... no chairs, 275 desks, and 100 tables... maximum profit contributation is $8,400..."


# Formulation

Descision variables

* (c)hair, (d)esk, (t)able

Objective function

* Maximize: 15c + 24d + 18t
  
Subject to:

* Fabrication: 4c + 6d + 2t <= 1850
* Assembly: 3c + 5d + 7t <= 2400
* Shipping: 3c + 2d + 4t <= 1500
* Demand Potential:
    * c <= 360
    * d <= 300
    * t <= 100

# R translation

Descision variables
```{r}
dv = c("c","d","t")
```

Objective function
```{r}
obj = c(15,24,18)
names(obj) = dv
```

* naming (`names(obj) = dv`) our objective function variables makes `lppp(...)` look nicer

Requirments
```{r}
req.fab = c(4,6,2)
req.ass = c(3,5,7)
req.shp = c(3,2,4)
req.dpc = c(1,0,0)
req.dpd = c(0,1,0)
req.dpt = c(0,0,1)

lhs = rbindPattern("req\\..+")
rm(req.fab, req.ass, req.shp, req.dpc, req.dpd, req.dpt)
```

* `lp` assumes all varables are >= 0, so we don't need to add that in too
* `req.???` is a special naming convention used by `rbindPattern` which will line up the left hand side (`lhs`) and right hand side (`rhs`) for us in an easy manner
* `req.dpc` is demand potential c <= 360
    * c <= 360 is rewriten as
    * 1c <=360 is rewriten as
    * 1c + 0d + 0t <= 360 is rewriten as
    * `c(1,0,0)`
* The regular expression `req\\..+` matches anything that starts with "req." and has at least one other character on the end

Constraints
```{r}
con.fab = 1850
con.ass = 2400
con.shp = 1500
con.dpc = 360
con.dpd = 300
con.dpt = 100

rhs = rbindPattern("con\\..+")
rm(con.fab, con.ass, con.shp, con.dpc, con.dpd, con.dpt)
```

Solve using `lpSolve`
```{r}
dir = rep_len("<=", dim(lhs)[1])

model = lpSolve::lp ("max", obj, lhs, dir, rhs)
model$objval
model$solution
```

* `model$objval` is the value of the __Objective Function__
* `model$solution` are the values of the __Descision Variables__
* This matches the book result of c=0,d=275,t=100 yealds 8400

Solve using `NlcOptim`
```{r}
startingpos = c(0,0,0)
objfun=function(x) {
  result = 15*x[1] + 24*x[2] + 18*x[3]
  return(-1*result)
}

lhs = rbind(lhs, nonzeroRequirment(dv))
rhs = rbind(rhs, nonzeroConstraint(dv))

model = NlcOptim::solnl(startingpos, objfun, A = lhs, B = rhs, maxIter = 4000)
-1*model$fn
as.vector(t(model$par))
```

* Technicaly `NlcOptim` only does minimum. To fix that make sure you `return(-1*result)`
* `lp` by default makes all the variables non-zero by default. `solnl` does not assume that, so we need to add in `nonzeroRequirment(...)` and `nonzeroConstraint(...)`
* `-1*model$fn` is the value of the __Objective Function__. It is the negitive because we fliped the result as above to get the "Maximum"
* `as.vector(t(model$par))` are the values of the __Descision Variables__
* This matches the book result of c=0,d=275,t=100 yealds 8400
