---
title: "9.2 Allocation Models"
author: "Mark Newman"
output:
  html_document: default
---

# Clear everything
```{r}
rm(list = ls())
options(scipen = 999)
```

# Required packages
```{r message=FALSE, warning=FALSE}
if (!require('lpSolve')) install.packages('lpSolve', quiet=TRUE)
source("../Helper Scripts/rbindPattern.r")
```

# LP Formulation

Descision variables

* (c)hair, (d)esk, (t)able

Maximize:

* 15c + 24d + 18t
  
Subject to:

* Fabrication: 4c + 6d + 2t <= 1850
* Assembly: 3c + 5d + 7t <= 2400
* Shipping: 3c + 2d + 4t <= 1500
* Demand Potential:
  * c <= 360
  * d <= 300
  * t <= 100

Objective function
```{r}
obj = c(15,24,18)
```

Requirments
```{r}
req.fab = c(4,6,2)
req.ass = c(3,5,7)
req.shp = c(3,2,4)
req.dpc = c(1,0,0)
req.dpd = c(0,1,0)
req.dpt = c(0,0,1)
```

* `lp` assumes varables >=0
* `req.dpc` is demand potential c <= 360.
  * c <= 360 is rewriten as
  * 1c <=360 is rewriten as
  * 1c + 0d + 0t <= 360 is rewriten as
  * `c(1,0,0)`

Constraints
```{r}
con.fab = 1850
con.ass = 2400
con.shp = 1500
con.dpc = 360
con.dpd = 300
con.dpt = 100
```

Solve (general LP)
```{r}
lhs = rbindPattern("req\\..+")
dir = rep_len("<=", dim(lhs)[1])
rhs = rbindPattern("con\\..+")

model = lp ("max", obj, lhs, dir, rhs)
model$objval
model$solution

cbind((rowSums(lhs*model$solution)), dir, rhs)
```

* `rbindpattern`
* `req\\..+` matches anything that starts with "req." and has at least one other character on the end
* `model$objval` is the value of the __Objective Function__
* `model$solution` are the values of the __Descision Variables__
* `cbind(...)` is there so you can _double check_ your work
