---
title: "8.5.1 Factory Location"
author: "Mark Newman"
output:
  html_document: default
---

# Clear everything
```{r}
rm(list = ls())
options(scipen = 999)
```

# Required packages
```{r message=FALSE, warning=FALSE}
if (!require('NlcOptim')) install.packages('NlcOptim', quiet=TRUE)
```

* `NlcOptim` is a popular R package implementing a nonlinear solver.

# Formulation

Descision variables

* location (x,y)

Objective function

* Minimize: sum(sqrt((x - x1)^2 + (y - y1)^2))
  
Subject to:

* None

# Tables
Site | x | y
:--:|:--:|:--:
 1 |   5 | 41
 2 |  20 | 10
 3 |  44 | 48
 4 |  60 | 58
 5 | 100 |  4
 6 | 138 | 80
 7 | 150 | 40
 8 | 170 | 18
 9 | 182 |  2
10 | 190 | 56


# R translation

Tables
```{r}
sites = matrix(c(1,5,41,2,20,10,3,44,48,4,60,58,5,100,4,6,138,80,7,150,40,8,170,18,9,182,2,10,190,56), nrow = 10, ncol = 3, byrow = TRUE)
colnames(sites) = c("Site", "X", "Y")
```

Objective function
```{r}
objfun = function(x) {
  
  print(as.vector(x))
  
  s = 0
  t = sites[,2:3]
  len = dim(t)[1]
  
  for(i in 1:len) {
    s = s + sqrt((t[i,1] - x[1])^2 + (t[i,2] - x[2])^2)
  }
  
  return(s)
}
abs(objfun(c(0,0)) - 1178.106) <= .001
```

* R has a funny notion of scoping. Everything is in scope, unless it is overwriten. In this case we don't pass in `sites`, but it is still available for us to do things with.
* The only reason `print(as.vector(x))` is in the function is so we can see the path we took to solve.
* **ALWAYS** double check the objective function, sometimes it _looks_ right, but is not. A single hand calculation goes a _lonnnng_ way.
    * `abs(...) <= .001` means we are the same rounded to 3 decimal places (Excel default)
    * `1178.106` comes from Excel

Linear Requirments / Constraints
```{r}
# This space left blank
```

* Requirments and Constraints are only added for linear equations. They are written in matrix form.

Nonlinear Requirments
```{r}
confun = function(x) { return(list(ceq = 0, c = 0)) }
```

* If you don't have any linear requirments, you **MUST** have a nonlinear requirment. `return(list(ceq = 0, c = 0))` is the No-Op constraint.
* ceq is for when you rewrite the nonlinear formula as xxx = 0.
* c is for when you can rewrite the nonlinear formula as xxx <= 0.
* You **MUST** rewrite the formulas like this. There is no option for left hand side / right hand side like in Linear Requirments.

Solve
```{r}
startingpos = c(0,0)

model1 = NlcOptim::solnl(startingpos, objfun=objfun, confun=confun, maxIter = 4000)
model1$fn
as.vector(t(model1$par))
```

* Starting at "0" is usualy a good idea to see what happens

Solve (better start)
```{r}
startingpos = c(mean(sites[,2]), mean(sites[,3]))

model2 = NlcOptim::solnl(startingpos, objfun=objfun, confun=confun, maxIter = 4000)
model2$fn
as.vector(t(model2$par))
```

* A "better" guess for `startingpos` is _somewhere_ inside the bounds.
    * We are doing `solnl(...)` 2 times juat as a sanity check.
* `model2$fn` is the value of the __Objective Function__.
* `as.vector(t(model2$par))` are the values of the __Descision Variables__

Compare results
```{r}
abs(model1$fn - model2$fn) <= .0001
abs(t(model1$par) - t(model2$par)) <= .0001
```

* x=114.0707, y=35.1722 yealds 670.638 is super close to the book result. and is identical if rounded to 1 decimal place.

# Book results

Per page 219 "The optimal solution ... is the location (114.1, 35.2) ... of approximately 670.6"