---
title: "8.5.2 Revenue Maximization"
author: "Mark Newman"
output:
  html_document: default
editor_options: 
  chunk_output_type: console
---

# Why choose this

Per page 219 "A common business problem involves maximizing revenue in the presence of a demand curve. A **demand curve** is simply a function that relates demand volumn to price. But because revenue is the product of price and volume, and volume depends on price, revenue is generally a nonlinear function of price"

# Formulation

Descision variables

$X = p_{d}, p_{e}$

* (D)aytime (P)rice, (E)vening (P)rice

Objective function

* Maxmize: $600d_{d}p_{d} + 840d_{e}p_{e}$

600 comes from minutes in the daytime period (8AM to 6PM). 840 comes from minutes in the evening period (6PM to 8AM).

Subject to

* (D)aytime (D)emand: $d_{d} = 600 - 5000p_{d} + 300p_{e}$
* (E)vening (D)emand: $d_{e} = 400 + 600p_{d} - 2500p_{e}$

# R translation

**Process**

* Clear everything
* Get required packages
* Add objective function
    * **ALWAYS** double check the objective function, sometimes it _looks_ right, but is not. A single hand calculation goes a _lonnnng_ way.
    * `nloptr` solves for minimum, to solve for max, use `return(-1*rev)` 
* Add in all constraints
    * We can infer the lower bounds because price will never be negitive
    * We can infer the upper bounds because demand will never be negitive
        * $0 = 600 - 5000p_{d} + 300*0 => p_{d} = 600/5000$
        * $0 = 400 + 600*0 - 2500p_{e} => p_{e} = 400/2500$
* Solve
    * When using a non-linear solver, always do at least 2 starting positions. Upper (`ub`) and lower (`lb`) bounds are usualy good bets.

**R Results**

$p_{d} = 0.07$, $p_{e} = 0.09$ yealds a revenue of $28043.85$.

**Book results**

Per page 221 "The optimal prices ... DP = 0.070 and EP = 0.091, with a maximum revenue of $28,044 per day"

* There is a misprint here somewhere. See **Code 2** below. DP = 0.070 + EP = 0.091 makes $28,042.98

**Code**
```{r results="hold"}
rm(list = ls())
options(scipen = 999)
if (!require('nloptr')) install.packages('nloptr', quiet=TRUE)
library(nloptr)
eval_f = function(x) {
  
  pd = x[1]
  pe = x[2]
  
  dd = 600 - 5000*pd + 300*pe
  de = 400 + 600*pd - 2500*pe
  
  rev = 600*dd*pd + 840*de*pe
  
  return(-1*rev)
}
eval_f(c(0,0)) == 0
lb = c(0,0)
ub = c(0.12,0.16)
sp1 = lb
sp2 = ub
opts = list(algorithm="NLOPT_LN_COBYLA", maxeval=4000, xtol_rel=1e-8)
model1 = nloptr(x0 = sp1, eval_f = eval_f, lb = lb, ub = ub, opts = opts)
model2 = nloptr(x0 = sp2, eval_f = eval_f, lb = lb, ub = ub, opts = opts)
round(-1* model1$objective, 2)
round(-1* model2$objective, 2)
round(model1$solution, 2)
round(model2$solution, 2)
print(model1)
print(model2)
```

**Code 2**
```{r results="hold"}
600*(600 - 5000*(0.070) + 300*(0.091))*(0.070) + 840*(400 + 600*(0.070) - 2500*(0.091))*(0.091)
```