---
title: "8.5.1 Factory Location"
author: "Mark Newman"
output:
  html_document: default
editor_options: 
  chunk_output_type: console
---

# Tables
Site | x | y
:--:|:--:|:--:
 1 |   5 | 41
 2 |  20 | 10
 3 |  44 | 48
 4 |  60 | 58
 5 | 100 |  4
 6 | 138 | 80
 7 | 150 | 40
 8 | 170 | 18
 9 | 182 |  2
10 | 190 | 56

# Formulation

Descision variables

* location (x,y)

Objective function

* Minimize: $\sum_{i=1}^{10}\sqrt{(x - x_{i})^2 + (y - y_{i})^2}$
  
Subject to

* None

# R translation

Clear everything
```{r}
rm(list = ls())
options(scipen = 999)
```

Required packages
```{r}
if (!require('nloptr')) install.packages('nloptr', quiet=TRUE)
library(nloptr)
```

Tables
```{r}
sites = matrix(c(1,5,41,2,20,10,3,44,48,4,60,58,5,100,4,6,138,80,7,150,40,8,170,18,9,182,2,10,190,56), nrow = 10, ncol = 3, byrow = TRUE)
colnames(sites) = c("Site", "X", "Y")
```

Objective function
```{r}
eval_f = function(x) {
  
  s = 0
  t = sites[,2:3]
  len = dim(t)[1]
  
  for(i in 1:len) {
    s = s + sqrt((t[i,1] - x[1])^2 + (t[i,2] - x[2])^2)
  }
  
  return(s)
}
abs(eval_f(c(0,0)) - 1178.106) <= .001
```

* R has a funny notion of scoping. Everything is in scope, unless it is overwriten. In this case we don't pass in `sites`, but it is still available for us to do things with.
* **ALWAYS** double check the objective function, sometimes it _looks_ right, but is not. A single hand calculation goes a _lonnnng_ way.
    * `abs(...) <= .001` means we are the same rounded to 3 decimal places (Excel default)
    * `1178.106` comes from Excel

Add in all constraints
```{r}
lb = c(min(sites[,2]), min(sites[,3]))
ub = c(max(sites[,2]), max(sites[,3]))
```

* Burried in this problem is a notion of centroid. Due to the way the formula is setup, we can infer bounds

Solve
```{r results="hold"}
sp1 = lb
sp2 = ub
opts = list(algorithm="NLOPT_LN_COBYLA", maxeval=4000)

model1 = nloptr(x0 = sp1, eval_f = eval_f, lb = lb, ub = ub, opts = opts)
model2 = nloptr(x0 = sp2, eval_f = eval_f, lb = lb, ub = ub, opts = opts)

round(model1$objective, 2)
round(model2$objective, 2)
round(model1$solution, 2)
round(model2$solution, 2)

print(model1)
print(model2)
```

* When using a non-linear solver, always do at least 2 starting positions. Upper (`ub`) and lower (`lb`) bounds are usualy good bets.
* x=114.07, y=35.17 yealds 670.64 is super close to the book result and is identical if rounded to 1 decimal place.

# Book results

Per page 219 "The optimal solution ... is the location (114.1, 35.2) ... of approximately 670.6"